---
title: "Replicate Experiment Update"
author: "Jeffrey Weidner"
format: 
  html: 
    toc: true
    toc-depth: 3
execute:
  echo: false
  messages:  false
  warning: false
  error: false
  
editor: visual
---

## Replicate Experiment Production Requirements Update

Documentation update for fun_msx.R to implement project team [specifications](https://app.box.com/s/u8ccadfsj2ix6do2a2eu5e8o1rmxyk4p) in box repository. This code and the associated files can be found in the [Replicate-Experiment](https://github.com/Jeff-Qualsci/Replicate-Experiment) repository. This script is used primarily on the server side of the Replicate Experiment web tool to perform the statistical analysis and generate results and statistics table for the web tool. It also includes examples for the figures using ggplot, though the web tool uses plotly to create interactive figures.

The updated specifications were generated to improve the ease of use/interpretation for users. The base functionality of the primary functions (`repexp.efficacy()` and `repexp.potency()`) is the same. Both functions use the same input and return a list of 5 objects (3 data tables and 2 ggplot figures) with the same object names as the previous version. However there are changes to the underlying code and helper functions to accommodate the updated requirements. Changes to column names are done as the final report list is created. The tidyverse allows the use of non-standard characters in object names, if they are enclosed by back ticks. These back ticks are automatically removed by tidvyerse aware functions such as readr::write_csv() or gt::gt(). The object names must be enclosed in back ticks if they are passed to base R functions.

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(patchwork)
library(gt)

source(file.path("R", 'fun_msx.R'))

reports_dir <- file.path("Reports")
```

## Replicate Experiment - Potency

The following illustrates the updated requirements for the Replicate Experiment in Potency mode.

### Bland-Altman figure

```{r}
#| label: potency_ba
#| echo: true 

tst_pot <- read_csv("Data/RepExpPotency.csv")

pot_report <- repexp.potency(tst_pot)

pot_report$MDPlot
```

Note change in figure title to "Potency Ratio vs Geometric Mean".

```{r}
#| echo: true

pot_report$Stats
```

The Stats element of the returned list now uses back ticks to enclose non-standard r object names. These back ticks must be included if passing these objects to base R functions which are not compliant with the standard. However when exporting with `gt::gt` or `readr::write_csv` these are automatically removed as shown below.

```{r}
#| label: potency_ba_stats
#| echo: true 

pot_ba_stats <- pot_report$Stats %>% 
  gt() %>% 
  tab_header(title = "Statistics (n = # of samples)")

pot_ba_stats

```

Column names and order as specified in the Box document.

```{r}
#| label: potency_MSRx
#| echo: true 

msrx_tbl <- pot_report$MSx_Tbl %>% 
  gt %>% 
  tab_header(title = "MSR of the Geometric Mean (N = # of inter-run replicates)")

msrx_tbl
```

**Rep-Exp Acceptance Criteria Text**

The following should be included in both the UI and exported report for the Potency mode of the Replicate-Experiment.

\
"The Replicate-Experiment MSR applies to potency results for samples that are tested once in a single experiment per sample. The MSR can be decreased significantly if every sample is tested 2 or 3 times, and the geomean potency is used to compare compounds. Both compounds being compared have to have at least “N” repeats each (i.e., “N” = the compound with fewest repeats). The MSR of Geometric Mean applies when compounds are compared using the geometric mean IC50 per compound (averaged across experiments). This calculation is only valid if the results are obtained from independent experiments and not replicates within the same experiment. The calculated MSR of Geometric Mean is the smallest ratio of the geometric mean IC50s between two compounds that is statistically discernable (a.k.a, statistically significant).”

### Correlation Plot

```{r}
#| label: potency_corr
#| echo: true 

pot_report$CorrPlot
```

Note the ggplot object is provided for reference. The axes have been switched as requested in the Box document.

## Efficacy

Similar changes were made to the output generated by `repexp.efficacy()`. There are no Acceptance Criteria in Efficacy mode to accompany the MSx_tbl because, unlike potency data, the variability of efficacy data is not usually constant across the dynamic range of the assay.
